Computing Vowel Elision in Yoruba Collocations

### 2023-05-07 paper.tex
Address peer review comments

To break the tie in \pref{tableau:onset-maxc-tie}, we introduce a morphemic
faithfulness constraint against vowel deletion at the beginnings of morpheme
boundaries. We call this constraint \maxplusv. Though this means that we must
consult morphology in order to determine correct phonological outputs,
observing the phenomenon in \pref{fig:deletion-simple}, we see that it is
fundamentally a morphological one, so we posit that relying on morphology here
presents the most productive path forward. We also propose its sibling
constraint, \maxvplus, which prohibits vowels from being deleted in
morpheme-final positions. We define both of these constraints in
\pref{def:maxvplus-and-maxplusv}.

@@

This modification enables us to account for the phenomena in
\pref{fig:deletion-except-2} by ranking /i/ as lower in the sonority scale of
Yoruba. This elegantly deals with the issue of /i/ deletion at the beginning of
the second vowel being preferred over /u/ deletion at the end of the first
vowel, because, as discussed before, a violation of \ssp{} represents a
maximally severe violation of the language's phonology. This is explicated in
the tableau shown here.

@@

As we can see here, the fact that morpheme-final [u] deletion is dispreferred
over morpheme-initial [i] deletion (which violates a fairly high-ranked
\maxplusv{} constraint) in the phenomena shown in \pref{fig:deletion-except-2}
is straightforwardly dealt with by taking the language-generic SSP ranking and
theorizing that Yoruba possesses a specialized variant of it, as many languages
do.

>>>>

To break the tie in \pref{tableau:onset-maxc-tie}, we introduce a morphemic
faithfulness constraint against vowel deletion at the beginnings of morpheme
boundaries. We call this constraint \maxplusv. Though this means that we must
consult morphology in order to determine correct phonological outputs,
observing the phenomenon in \pref{fig:deletion-simple}, we see that it is
fundamentally a morphological one, so we posit that relying on morphology here
presents the most productive path forward. We also propose its sibling
constraint, \maxvplus, which prohibits vowels from being deleted in
morpheme-final positions. We define both of these constraints in
\pref{def:maxvplus-and-maxplusv}. Note that we could define these without
specifying \textsc{-V}; that is, they could simply regulate against deletion
at morpheme boundaries, but we have chosen to specialize these to our analysis
to avoid unintended interactions with deletion in Yoruba unrelated to our work.

@@

This modification enables us to account for the phenomena in
\pref{fig:deletion-except-2} by ranking /i/ as lower in the sonority scale of
Yoruba. This deals with the issue of /i/ deletion at the beginning of
the second vowel being preferred over /u/ deletion at the end of the first
vowel, because, as discussed before, a violation of \ssp{} represents a
maximally severe violation of the language's phonology. This is explicated in
the tableau shown here.

@@

As we can see here, the fact that morpheme-final [u] deletion is dispreferred
over morpheme-initial [i] deletion (which violates a fairly high-ranked
\maxplusv{} constraint) in the phenomena shown in \pref{fig:deletion-except-2}
is straightforwardly dealt with by taking the language-generic SSP ranking and
theorizing that Yoruba possesses a specialized variant of it, as many languages
do. This sonority scale is motivated only linguistically, within the context
of our analysis, and not psycho-acoustically-- though that would be an
interesting direction for future work.
